Behavioral Analysis of the Agent-Based Community Grid Solution for the Large Hadron Collider beauty Experiment

### 2011-11-12 paper_update.tex
halfway there

\usepackage{url}
\usepackage{graphicx}
\usepackage{cite}
%\usepackage[center]{caption}
\usepackage{authblk}
\begin{document}
\title{Behavioral Analysis of the Agent-Based Community Grid Solution for the Large Hadron Collider beauty Experiment}
\author[1,3]{Daniela Remenska}
\author[2]{Tim A.C. Willemse}
\author[1]{Henri Bal}
\author[1]{Kees Verstoep}
\author[1]{Wan Fokkink}
\author[3]{Jeff Templon}

@@

The paper is organized as follows. Section 2 introduces the architecture
of DIRAC, focusing on the two studied subsystems. Section 3 gives a brief
overview of the mCRL2 language, and describes our approach to abstracting
and modeling the behavior of these subsystems. Section 4 presents the
analysis with the mCRL2 toolset and the issues detected. Section 5
concludes and discusses future work.

@@

@@

The driving force of DIRAC is the Workload Management System (WMS). Taking into
account that the distributed computing environment is intrinsically unstable, a
\textit{Pilot Job paradigm}, illustrated in Fig.~\ref{fig:DIRAC-WMS}, was chosen as an efficient way to
implement a pull scheduling mechanism.  Pilot jobs are simply resource
reservation processes without an actual payload defined apriori. They are
submitted to the Grid worker nodes with the aim of checking the sanity of the
operational environment just before pulling and executing the real payload. This
hides the fragility of the underlying resources and increases the job success
rate, from the perspective of end users Centralized \textit{Task Queues} contain all the
pending jobs, organized in groups according to their requirements. This enables
efficient application of job priorities, enforcing faire-share policies and
quotas, which are decided within the LHCb community, without the need of any
effort from the Grid sites.  
\begin{figure}[t]
\includegraphics[width=\linewidth,keepaspectratio=true]{./DIRAC_WMS.png}
\centering
\caption{DIRAC Workload Management System \cite{DIRAC_pilot_WMS}}
\label{fig:DIRAC-WMS}
\end{figure}
 \begin{figure}[b]
\includegraphics[width=\linewidth,keepaspectratio=true]{./dirac-primary-states.png}
\centering
\caption{Job state machine \cite{ProductionShifterGuide}}
\label{fig:DIRAC-job-state-machine}
\end{figure}

The basic flowchart describing the evolution of a job’s states is depicted in
Fig.~\ref{fig:DIRAC-job-state-machine}. After submission through the \textit{Job Manager} (“Received”), the complete
job description is placed in the DIRAC job repository (the Job DB). Before jobs
become eligible for execution, a chain of optimizer agents check and prioritize
them in queues, utilizing the parameters information from the Job DB. \textit{The Job
Sanity Agent} can meaningfully fail jobs with impossible requirements, and
prevent unnecessary submission to the Grid resources.The \textit{Input Data Agent}
performs a resolution of possible target computing resources based on the input
data requirements. If the requested data resides on tape storage, the \textit{Job
Scheduling Agent} will pass the control to a specialized \textit{Stager} service (part of
the SMS explained in the next section), before placing the job a Task Queue
(“Waiting”). Based on the complete list of pending payloads, a specialized \textit{Task
Queue Director} submits pilots to the computing resources via the gLite WMS. After
a \textit{Matcher} service pulls the most suitable payload for a pilot (“Matched”), a \textit{Job
Wrapper} object is created on the worker node, responsible for retrieving the
input sandbox, performing software availability checks, executing the actual
payload on the worker node (“Running”), and finally uploading any output data
necessary (“Done” or “Failed”). The wrapper can catch any failure exit state
reported by the running physics applications. At the same time, a \textit{Watchdog}
process is instantiated to monitor the behavior of the Job Wrapper and send
heartbeat signals to the monitoring service. It can also take actions in case
resources are soon to be exhausted, the payload stalls, or a management command
for killing the payload is received.

@@

A simplified view of the system is shown in  Fig.~\ref{fig:DIRAC-SMS}. The workflow is initiated
with the Job Scheduling agent detecting that a job is assigned to process files
only available on tape storage. It sends a request for staging (i.e., creating a
cached replica) to the \textit{Storage Manager Handler} service along with the
list of files and a callback method to be invoked when the request has been
processed. The Storage Management DB is immediately populated with records which
are processed by a sequence of agents in an organized fashion. The relevant
tables in the SMS DB are the \textit{Tasks} and \textit{CacheReplicas}, whose
entities maintain a state observed and updated by these agents. Tasks maintain
general information about every job requesting a service from the SMS. The
details about every file (i.e., the Storage Element (SE) where it resides, the
size, checksum, number of tasks that requested it), are kept in the CacheReplicas
table. Other auxiliary tables maintain the relationship between these entities.
\begin{figure}[t]
\includegraphics[width=0.9\linewidth,keepaspectratio=true]{./SMS_simplified.png}
\centering
\caption{DIRAC Storage Management System}
\label{fig:DIRAC-SMS}
\end{figure}
The processing begins with the \textit{Request Preparation Agent}. It selects all the
“New” entries for files, checks whether they are registered in a special logical
file catalog, and retrieves their metadata. In case of problematic catalog
entries, it can update the status of the CacheReplicas and the related Tasks
entries to “Failed”. For non-problematic files it carries out transition to
“Waiting”. The \textit{StageRequest Agent} is responsible for placing the actual staging
requests for all “Waiting” entries, via dedicated storage middleware that
communicates with the tape backends. These requests are grouped by SE prior to
submission, and carry information about the requested (pin) lifetime of the
replicas to be cached. If certain pathologies are discovered (i.e., lost,
unavailable or zero-sized files on tape), it can update the corresponding
entries to “Failed” in a similar manner. Otherwise, they are promoted to
“StageSubmitted”. The agent responsible for monitoring the status of submitted
requests is the \textit{StageMonitor Agent}. It achieves this by interrogating the
storage middleware to see if the “StageSubmitted” files are successfully
replicated on disk cache. In case of success, the CacheReplicas and their
corresponding Tasks entries are updated to “Staged”. Various circumstances of
tape or middleware misbehavior can also fail the staging requests. The last one
in the chain is the \textit{RequestFinalization Agent}. The tasks which are in their
final states (“Staged” or “Failed”) are cleared from the database, and callbacks
are performed to the WMS, which effectively wakes up the corresponding jobs. If
there are no more associated tasks for particular replicas, the respective
CacheReplicas entries are also removed.

@@

>>>>

\usepackage{color}
\usepackage{url}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{listings}
\definecolor{light-gray}{gray}{0.80}
%\usepackage[center]{caption}
\usepackage{authblk}
\begin{document}
\title{Behavioral Analysis of the Agent-Based Community Grid Solution for the Large Hadron Collider beauty Experiment}
\author[1,3]{Daniela Remenska}
\author[2]{Tim A.C. Willemse}
\author[1]{Henri Bal}
\author[1]{Kees Verstoep}
\author[1]{Wan Fokkink}
\author[3]{Jeff Templon}

@@

The paper is organized as follows. Section 2 introduces the architecture
of DIRAC, focusing on the two subsystems chosen as case studies. Section 3 gives a brief
overview of the mCRL2 language, and describes our approach to abstracting
and modeling the behavior of these subsystems. Section 4 presents the
analysis with the mCRL2 toolset and the issues detected. Section 5
concludes and discusses future work.

@@

!! Mention that there are several subsystems part of DIRAC's architecture.
We focus on two related subsystems that were considered essential, yet 
are the ones where problematic state changes were most often encountered, 
and difficult to trace and correct.

@@

The driving force of DIRAC is the Workload Management System (WMS). Taking into
account that the distributed computing environment is intrinsically unstable, a
\textit{Pilot Job paradigm}, illustrated in Fig.~\ref{fig:DIRAC-WMS}, was chosen as an efficient way to
implement a pull scheduling mechanism.  Pilot jobs are simply resource
reservation processes without an actual payload defined apriori. They are
submitted to the Grid worker nodes with the aim of checking the sanity of the
operational environment just before pulling and executing the real payload. This
hides the fragility of the underlying resources and increases the job success
rate, from the perspective of end users Centralized \textit{Task Queues} contain all the
pending jobs, organized in groups according to their requirements. This enables
efficient application of job priorities, enforcing faire-share policies and
quotas, which are decided within the LHCb community, without the need of any
effort from the Grid sites.  
\begin{figure}[t]
\includegraphics[width=\linewidth,keepaspectratio=true]{./DIRAC_WMS1.png}
\centering
\caption{DIRAC Workload Management System \cite{DIRAC_pilot_WMS}}
\label{fig:DIRAC-WMS}
\end{figure}
 \begin{figure}[b]
\includegraphics[width=\linewidth,keepaspectratio=true]{./dirac-primary-states.png}
\centering
\caption{Job state machine \cite{ProductionShifterGuide}}
\label{fig:DIRAC-job-state-machine}
\end{figure}

The basic flowchart describing the evolution of a job’s states is depicted in
Fig.~\ref{fig:DIRAC-job-state-machine}. After submission through the \textit{Job Manager} (``Received''), the complete
job description is placed in the DIRAC job repository (the Job DB). Before jobs
become eligible for execution, a chain of optimizer agents check and prioritize
them in queues, utilizing the parameters information from the Job DB. \textit{The Job
Sanity Agent} can meaningfully fail jobs with impossible requirements, and
prevent unnecessary submission to the Grid resources.The \textit{Input Data Agent}
performs a resolution of possible target computing resources based on the input
data requirements. If the requested data resides on tape storage, the \textit{Job
Scheduling Agent} will pass the control to a specialized \textit{Stager} service (part of
the SMS explained in the next section), before placing the job a Task Queue
(``Waiting''). Based on the complete list of pending payloads, a specialized \textit{Task
Queue Director} submits pilots to the computing resources via the gLite WMS. After
a \textit{Matcher} service pulls the most suitable payload for a pilot (``Matched''), a \textit{Job
Wrapper} object is created on the worker node, responsible for retrieving the
input sandbox, performing software availability checks, executing the actual
payload on the worker node (``Running''), and finally uploading any output data
necessary (“Done” or ``Failed''). The wrapper can catch any failure exit state
reported by the running physics applications. At the same time, a \textit{Watchdog}
process is instantiated to monitor the behavior of the Job Wrapper and send
heartbeat signals to the monitoring service. It can also take actions in case
resources are soon to be exhausted, the payload stalls, or a management command
for killing the payload is received.

@@

A simplified view of the system is shown in  Fig.~\ref{fig:DIRAC-SMS}. The workflow is initiated
with the Job Scheduling agent detecting that a job is assigned to process files
only available on tape storage. It sends a request for staging (i.e., creating a
cached replica) to the \textit{Storage Manager Handler} service along with the
list of files and a callback method to be invoked when the request has been
processed. The Storage Management DB is immediately populated with records which
are processed by a sequence of agents in an organized fashion. The relevant
tables in the SMS DB are the \textit{Tasks} and \textit{CacheReplicas}, whose
entities maintain a state observed and updated by these agents. Tasks maintain
general information about every job requesting a service from the SMS. The
details about every file (i.e., the Storage Element where it resides, the
size, checksum, number of tasks that requested it), are kept in the CacheReplicas
table. Other auxiliary tables maintain the relationship between these entities.
\begin{figure}[t]
\includegraphics[width=0.9\linewidth,keepaspectratio=true]{./SMS_simplified.png}
\centering
\caption{DIRAC Storage Management System}
\label{fig:DIRAC-SMS}
\end{figure}

The processing begins with the \textit{Request Preparation Agent}. It selects all the
``New'' entries for files, checks whether they are registered in a special logical
file catalog, and retrieves their metadata. In case of problematic catalog
entries, it can update the status of the CacheReplicas and the related Tasks
entries to ``Failed''. For non-problematic files it carries out transition to
``Waiting''. The \textit{StageRequest Agent} is responsible for placing the actual staging
requests for all ``Waiting'' entries, via dedicated storage middleware that
communicates with the tape backends. These requests are grouped by SE prior to
submission, and carry information about the requested (pin) lifetime of the
replicas to be cached. If certain pathologies are discovered (i.e., lost,
unavailable or zero-sized files on tape), it can update the corresponding
entries to ``Failed'' in a similar manner. Otherwise, they are promoted to
``StageSubmitted''. The agent responsible for monitoring the status of submitted
requests is the \textit{StageMonitor Agent}. It achieves this by interrogating the
storage middleware to see if the ``StageSubmitted'' files are successfully
replicated on disk cache. In case of success, the CacheReplicas and their
corresponding Tasks entries are updated to ``Staged''. Various circumstances of
tape or middleware misbehavior can also fail the staging requests. The last one
in the chain is the \textit{RequestFinalization Agent}. The tasks which are in their
final states (``Staged'' or ``Failed'') are cleared from the database, and callbacks
are performed to the WMS, which effectively wakes up the corresponding jobs. If
there are no more associated tasks for particular replicas, the respective
CacheReplicas entries are also removed.

@@

\section{System modeling}
\subsection{The mCRL2 language and toolset}
tbw
\subsection{From DIRAC to mCRL2}

Any formal analysis uses a simplified description of the real system. Even in the 
best possible scenario, where the target implementation language and the modeling 
notation are very close (\cite{Java_PathFinder},\cite{Musuvathi04modelchecking}), 
it is practically impossible to avoid the use of abstraction to create a simplified model. 
Software implementations are large and often contain many details that can be irrelevant 
for the intended analysis. Abstraction aims at reducing
the program's state space in order to overcome the resource limitations \cite{Pelánek08fightingstate} encountered during model-checking.
Furthermore, specification languages describe \textit{what} is being done, 
abstracting away from the details of \textit{how} things are done. 
Then, the ultimate question is: \textit{how do we establish correspondence between model and code?}

High-level design documents are a good starting point, but insufficient for 
building a sound model.
In absence of more detailed up-to-date behavioral models, 
we based our models on the source code and discussions with developers.
A popular abstraction technique
for identifying code subsets (slices) that (potentially) affect particular 
variables of interest (slicing criteria) is program slicing \cite{Hatcliff99slicingsoftware}. 
It reduces the behavior of a program
by removing control statements and data structures deemed irrelevant for 
the criteria. Unfortunatelly, to the best of our knowledge, the only practical
applications of automated slicing have been
on C/C++ and Java programs.[references] Research has not yet
matured on this topic for dynamic languages, such as Python. 
Therefore, we performed the program slicing manually, relying on the Eclipse IDE for 
reverse-engineering and dependency analysis of the subsystems.

Given the complexity of the reccurent invalid state transitions of entities within DIRAC, 
we considered the possible race conditions caused by multiple agents updating 
the service states to be the target of our analysis. We limit the scope
to analysis of the following entities: \textit{Tasks} (SMS), \textit{CacheReplicas} (SMS) and \textit{Jobs} (WMS).
This determines our slicing criteria. In the following, we use the SMS as a case study
for describing the established modeling guidelines. They can be 
diractly applied to other DIRAC systems.

As already explained, all agents repeat the same logic in every subsequent iteration:first they
read some entries of interest from the service database, then they process the cached data, and finally
they may write or update entries back to the central database, based on decisions from the processing step.
The frontend for all interractions with the storage are the services, essentially acting as wrappers around it.
Take as an example the code snippet\footnote{This method is $\sim$100 lines of code in reality} 
in Listing~\ref{listing1}, 
from the simplest SMS agent: the \textit{Stage Request Agent}. 
Although much of the code is omitted for clarity, care is taken to preserve
the parts necessary for ilustrating the basic idea.
 (and yet it looks baffling!)

\definecolor{light-gray}{gray}{0.90}

\lstset{language=Python,tabsize=2,caption={Request Preparation Agent },
  stringstyle=\ttfamily,
  %basicstyle=\small,
  label=listing1,
  %frame=trbl,
  breaklines=true,
  showstringspaces=false,
  boxpos=t,
  belowcaptionskip=5pt
 }

\begin{lstlisting}[float=tp,escapechar=!,basicstyle=\ttfamily\fontsize{7}{7}\selectfont]
  def prepareNewReplicas( self ):
    """ This is the first logical task to be executed 
    """ and manages the New->Waiting transition of the Replicas
    res = self.__getNewReplicas()
    !\colorbox{light-gray}{res = self.getNewReplicas()}!
    if not res['Value']:
      gLogger.info('There were no New replicas found')
      return res
    [...]
    # Obtain the replicas from the FileCatalog
    res = self.__getFileReplicas( fileSizes.keys() )
    if not res['OK']:
      return res
    failed.update( res['Value']['Failed'] )
    terminal = res['Value']['ZeroReplicas']
    fileReplicas = res['Value']['Replicas']
    [...]
    # Check the replicas exist at the requested site
    replicaMetadata = []
    for lfn, requestedSEs in replicas.items():
      lfnReplicas = fileReplicas[lfn]
      for requestedSE, replicaID in requestedSEs.items():
        if not requestedSE in lfnReplicas.keys():
          terminalReplicaIDs[replicaID] = "LFN not registered at requested SE"
          replicas[lfn].pop(requestedSE)
        else:
          replicaMetadata.append((replicaID, lfnReplicas[requestedSE], fileSizes[lfn]))

    # Update the states of the files in the database
    if terminalReplicaIDs:
      gLogger.info('%s replicas are terminally failed.' % len( terminalReplicaIDs ) )
      !\colorbox{light-gray}{res = self.stagerClient.}!
	  !\colorbox{light-gray}{updateReplicaFailure(terminalReplicaIDs)}!
    if replicaMetadata:
      gLogger.info('%s replica metadata to be updated.' % len( replicaMetadata ) )
      # Sets the Status='Waiting' of CacheReplicas records 
      #that are OK with catalogue checks
      !\colorbox{light-gray}{res = self.stagerClient.}!
      !\colorbox{light-gray}{updateReplicaInformation( replicaMetadata )}!
    return S_OK()

\end{lstlisting}

\section{Analysis and Issues}
\label{sec:Section_4}
\section{Conclusions and Future Work}
\label{sec:Section_5}
